Recent demography drives changes in linked selection across the maize genome

### 2015-11-24 NatGen/manuscript.tex
updated abstract and discussion

\subsection*{Demography of domestication} %JRI
Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.  We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites. This implies that genic polymorphism data are unable to tell the complete or accurate demographic history of maize, but the rapid recovery of diversity outside of genes demonstrates that sites far from genes can be reasonably used for demographic inference. Furthermore, by utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without making assumptions about its duration.

@@

\subsection*{Hard sweeps do not shape genome-wide diversity in maize} %JRI

Our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of purifying selection  \cite{enard2014}. 
Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability. 
Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}. 
Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{van2012historical, beissinger2014} all support the importance of standing variation as primary sources of adaptive variation. 
Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013}. Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.]{studer2011}.
Indeed, many adaptive traits in both maize \cite{wallace2014} and teosinte \cite{weber2008} are highly quantitative,  and adaptation in both maize \cite{hufford2012} and teosinte \cite{pyhajarvi2013} has involved selection on regulatory variation.

@@

\subsection*{Demography influences the efficiency of purifying selection} %JRI 
One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
We observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), but the opposite trend when we evaluate diversity using singleton polymorphisms (Figure \ref{fig:purify}B).
The efficiency of purifying selection is proportional to effective population size \cite{kimura1984}, and these results are thus consistent with our demographic analyses which show a domestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} followed by recent rapid expansion and a much larger modern $N_e$.

@@

\section*{Rapid changes in linked selection} %JRI 
Our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
While a number of authors have tested for selection using methods that explicitly incorporate or are robust to demographic change \cite{eyre2009estimating, chen2010, zeng2010} and others have compared estimates of the efficiency of adaptive and purifying selection across species \cite{popadin2012} or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size \cite{corbett2015,leffler2012}.
Our results show that demographic change over short periods of time can quickly change the dynamics of linked selection: mutations arising in extant maize populations are much more strongly impacted by the effects of selection on linked sites than would be suggested by analyses using long-term effective population size.
As many natural and domesticated populations have undergone considerable demographic change in their recent past, long-term comparisons of $N_e$ are likely not informative about current processes affecting allele frequency trajectories.

>>>>

%\subsection*{Demography of domestication} %JRI
Although a number of authors have investigated the demography of maize domestication \cite{eyre1998, tenaillon2004, wright2005}, these  efforts relied on data only from genic regions of the genome and made a number of limiting assumptions about the demographic model.  We show that diversity within genes has been strongly reduced by the effects of linked selection, such that even synonymous polymorphisms in genes are not representative of diversity at unconstrained sites. This implies that genic polymorphism data are unable to tell the complete or accurate demographic history of maize, but the rapid recovery of diversity outside of genes demonstrates that sites far from genes can be reasonably used for demographic inference. Furthermore, by utilizing the full joint SFS, we are able to estimate population growth, gene flow, and the strength of the domestication bottleneck without making assumptions about its duration.

@@

%\subsection*{Hard sweeps do not shape genome-wide diversity in maize} %JRI

Irrespective of ancient or current population size, our findings demonstrate that classic hard selective sweeps have not contributed substantially to genome-wide patterns of diversity in maize, a result we show is robust to concerns about power due to the effects of purifying selection  \cite{enard2014}. 
Although our approach ignores the potential for hard sweeps in noncoding regions of the genome, a growing body of evidence argues against hard sweeps as the prevalent mode of selection shaping maize variability. 
Among well-characterized domestication loci, only the gene \emph{tga1} shows evidence of a hard sweep on a missense mutation \cite{wang2015}, while several loci are consistent with ``soft sweeps''  from standing variation \cite{studer2011,gallavotti2004role} or multiple mutations \cite{wills2013}. 
Moreover, genome-wide studies of domestication \cite{hufford2012},  local adaptation \cite{Takuno15062015} and modern breeding \cite{van2012historical, beissinger2014} all support the importance of standing variation as primary sources of adaptive variation. 
Soft sweeps are expected to be common when $2N_e\mu_b \ge 1$, where $\mu_b$ is the mutation rate of beneficial alleles with selection coefficient $s_b$ \cite{messer2013}. Assuming a mutation rate of $3 \times 10^{-8}$ \cite{clark2005} and that on the order of $\approx 1-5\%$ of mutations are beneficial \cite{eyre2007}, this implies that soft sweeps should be common in both maize and teosinte for mutational targets $>>10kb$ --- a plausible size for quantitative traits or for regulatory evolution targeting genes with large up- or down-stream control regions \cite[e.g.]{studer2011}.
Indeed, many adaptive traits in both maize \cite{wallace2014} and teosinte \cite{weber2008} are highly quantitative,  and adaptation in both maize \cite{hufford2012} and teosinte \cite{pyhajarvi2013} has involved selection on regulatory variation.

@@

%\subsection*{Demography influences the efficiency of purifying selection} %JRI 
One of our more striking findings is that the impact of purifying selection on maize and teosinte qualitatively changed over time. 
We observe a more pronounced decrease in $\pi$ around genes in teosinte than maize (Figure \ref{fig:purify}A), but the opposite trend when we evaluate diversity using singleton polymorphisms (Figure \ref{fig:purify}B).
The efficiency of purifying selection is proportional to effective population size \cite{kimura1984}, and these results are thus consistent with our demographic analyses which show a domestication bottleneck and smaller long-term $N_e$ in maize \cite{eyre1998, tenaillon2004, wright2005, ross2009} followed by recent rapid expansion and a much larger modern $N_e$.

@@

%\section*{Rapid changes in linked selection} %JRI
Most importantly, our results demonstrate that consideration of long-term differences in $N_e$ cannot fully capture the dynamic relationship between demography and selection. 
While a number of authors have tested for selection using methods that explicitly incorporate or are robust to demographic change \cite{eyre2009estimating, chen2010, zeng2010} and others have compared estimates of the efficiency of adaptive and purifying selection across species \cite{popadin2012} or populations \cite{Elyashiv:2010ga}, previous analyses of the impact of linked selection on genome-wide diversity have relied on single estimates of the effective population size \cite{corbett2015,leffler2012}.
Our results show that demographic change over short periods of time can quickly change the dynamics of linked selection: mutations arising in extant maize populations are much more strongly impacted by the effects of selection on linked sites than would be suggested by analyses using long-term effective population size.
As many natural and domesticated populations have undergone considerable demographic change in their recent past, long-term comparisons of $N_e$ are likely not informative about current processes affecting allele frequency trajectories.
